\documentclass[prc]{revtex4}
\usepackage[dvips]{graphicx}
\usepackage{mathrsfs}
\usepackage{amsfonts}
\usepackage{lscape}

\usepackage{epic,eepic}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[dvips]{epsfig}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{bezier}
\usepackage{pstricks}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{braket}
\usepackage[dvips]{graphicx}
\usepackage{pst-plot}

\newcommand{\One}{\hat{\mathbf{1}}}
\newcommand{\eff}{\text{eff}}
\newcommand{\Heff}{\hat{H}_\text{eff}}
\newcommand{\Veff}{\hat{V}_\text{eff}}
\newcommand{\braket}[1]{\langle#1\rangle}
\newcommand{\Span}{\operatorname{sp}}
\newcommand{\tr}{\operatorname{trace}}
\newcommand{\diag}{\operatorname{diag}}
\newcommand{\bra}[1]{\left\langle #1 \right|}
\newcommand{\ket}[1]{\left| #1 \right\rangle}
\newcommand{\element}[3]
    {\bra{#1}#2\ket{#3}}

\newcommand{\normord}[1]{
    \left\{#1\right\}
}



\usepackage{amsmath}
\begin{document}


\title{Project for the final oral examination PHY981}
%\author{}
\maketitle
\section*{Studies of spectroscopic factos in Ar isotopes and $^{48}$Ca}

The aim of this project is to study the structure of selected low-lying states of the Ar isotopes and 
pertinent spectroscopic factors using Alex Brown's Nushellx code. 
There is considerable interest in neutron-rich nuclei between $N = 14$ and $N = 28$,
motivated particularly by the development of collectivity in nuclei such as
$^{40,42}$S and by questions
concerning the strength of the N = 28 shell gap far from stability (see Refs.~[1-5] and references
therein). The Ar isotopes from $^{34}$Ar ($N = 16$) to $^{46}$Ar ($N = 28$) are important for studying this
evolution of collectivity and for testing shell model calculations for neutron-rich nuclei in the
region. Because they have two proton holes in the $sd$ shell, the neutron-rich Ar isotopes offer a
relatively ‘simple’ case to probe proton-neutron interactions and build up a
microscopic description of nuclear structure in neutron rich nuclei near $N = 28$. Being close to $^{48}$Ca, there 
is great interest in understanding the role of such interactions.   
The lightest Ar isotopes can be studied with effective interactions for $sd$-shell only as a first trial. 
For the heavier ones, one would need to test 
effective interactions that include the $0f_{7/2}$ orbit as well and possibly the $1p_{3/2}$ single-particle state. 
In Refs.~[2,3], shell-model calculations where performed
with effective interactions that included parts of the $pf$-shell as well. 

The task here is to compute low-lying states of the Ar isotopes from
$N=16$ to $N=28$ and compute spectroscopic factors for the ground states and selected excited states
and compare with available data. To achieve this
you will need to use an effective interaction designed for the $1s0d$
and $1s0d1p0f$ shells. Since a full calculation in two major shells
may be beyond reach, you will need to truncate the number of particles
which can leave/occupy selected single-particle states. In the file
which contains the single-particle data, you should block some of the
single-particle orbis in the $1p0f$ shells.

The possible tasks are as follows
\begin{enumerate}
\item Compute the spectra  of the low-lying spectrum of the above even Ar isotopes using the $1s0d$ shell as degree of freedom first. Add thereafter the $0f_{7/2}$ single-particle orbit and study the behavior of your results. Compare with Refs.~[2,3]. 
\item Add thereafter the $1p_{3/2}$ single-particle state and commnet your results.
\item Compute thereafter the spectroscopic factors for the ground states
and compare with available data using the above model spaces. 
\end{enumerate}



Here follows a list of possible references.

\begin{enumerate}
\item  H. Scheit et al., Phys Rev Lett 77, 3967 (1996) 
\item  J. Retamosa et al., Phys Rev C 55, 1266 (1997) 
\item  S. Nummela et al., Phys Rev C 63, 044316 (2001) 
\item A.D. Davies et al., Phys Rev Lett 96, 112503 (2006)
\item  A.E. Stuchbery et al., Phys Rev C 74, 054307 (2006)
\end{enumerate}
\end{document}










